### 2018-06-29 Thesis/Parts/Data/Data.tex
Updated thesis

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
\header{Corpus}
The dataset we used for this research consists of articles published in 2017 which have been published in a journal that has, in 2017, atleast 150 publications. This results in a total dataset of $1.391.543$ articles from $3.759$ journals. Details about the corpus can be found in table~\ref{table:corpusSize}.
\begin{table}[hbt]
\begin{center}
\begin{tabular}{|l|c|c|c|}
\hline
 & Total count & Unique count & Average length \\
\hline\hline
Title words & $18.822.399$ & $939.665$ & 13,53  \\
\hline
Title tokens & $14.742.192$ & $230.805$ & 10,64 \\
\hline\hline
Abstract words & $264.653.020$ & $5.853.077$  & 190,19  \\
\hline
Abstract tokens & $171.474.473$ & $738.961$ & 123,71 \\
\hline\hline
Total words & $283.475.419$ & $6.209.769$  & 203,71 \\
\hline
Total tokens & $186.962.354$ & $763.475$ & 134,36 \\
\hline
\end{tabular}
\end{center}
\caption{Corpus size}\label{table:corpusSize}
\end{table}\\
\subheader{Text properties}
\citet{wiegand2018word} state that the Pareto distribution offers a good fit to the word occurrences in natural language. Their models show that, due to the additional parameters in the Pareto-III, the tail of the data fits better with the model than the Zipf model. This shows the relation between the word occurrences rank and the actual occurrences. This kind of word-occurrences distribution holds for many texts, including the writings of William Shakespeare and scientific texts and novels\cite{thurner2015understanding}.
The word occurrences in our corpus also follow the pattern of a Pareto distribution as described by~\citet{wiegand2018word}. Our distribution is visualized in figure~\ref{figure:wordTokenOccurrence}, which displays the occurrences of the first 500 tokens of the corpus.
\begin{figure}[hbt]
\includegraphics[width=6.5in]{Plots/word_occurrences}
\caption{Word and token occurrences}\label{figure:wordTokenOccurrence}
\end{figure}

\header{Datasets}
For this researched we used a (pre-made) tokenized dataset, created from the earlier described corpus, which reduces the total amount of words by 34\% (see table~\ref{table:corpusSize}). From this tokenized set, we created the embeddings and the TF-IDF feature vectors.
\subheader{Tokenization}
The following steps have been applied to the words to create a tokenized set:
\begin{enumerate}
\item{Removed punctuation}
\item{Removed all non-ASCII characters}
\item{Transformed all characters to lower-case}
\item{Removed stop-words, as provided by the NLTK\footnote{Natural Language ToolKit, \url{https://www.nltk.org/}} library}
\item{Removed numbers}
\item{Stemmed all words, using the stemmer provided by the NLTK library}
\end{enumerate}
These transformations reduced our dataset by 34\%, resulting in a tokenized set of $186.962.354$ tokens.

@@

@@

\header{Research environment}
The research will be done in a python-databricks environment, which uses spark, a library that offers tools to work with big-data. \citet{armbrust2015spark} state that Spark SQL lets programmers leverage the benefits of relational processing and lets SQL users call complex analytic libraries in Spark. This allows for much tighter integration between relational and procedural processing. The paper further states that Spark SQL makes it significantly simpler and more efficient to write data pipelines that mix relational and procedural processing while offering substantial speedups over previous SQL-on-Spark engines.

\header{Pipeline}
We procesessed the TF-IDF sets and the embedding sets via the same pipeline, using their common vector properties. This ensures comparable results, the pipeline is set-up as follows:
\begin{enumerate}
\item{Create training and validation set}
\item{Create journal embeddings}
\item{Categorize validation articles}
\item{Calculate performance metrics}
\end{enumerate}

@@

To calculate the distance between to vectors, cosine similarity is commonly used. We validated the quality of cosine similarity as a distance metrics by comparing it to all other similarity matrices available in the SciPy library, which we used to calculate the distances. We calculate the similarities based on the normalized embeddings, and compared the distance metrics based on the default embedding set. Table~\ref{table:distanceMetrics} shows the results of this validation.\\

@@

\subheader{Performance measurement}
We use multiple metrics to validate the performance of the embedding sets and TF-IDF sets on the categorization task. These metrics are:
\begin{enumerate}
\item{F1-score}
\item{Median \& average rank}
\item{Rank distribution}
\end{enumerate}
\subsubheader{F1-score}
We define the positive \& negative metrics as follows:\\
\begin{jumpin}
$True Positive = $ Articles that are correctly matched to the current journal\\
$False Positive = $ Articles that are incorrectly matched to other journals\\
$False Negative = $ Articles that are incorrectly matched to the current journal\\
\end{jumpin}
We used these metrics to calculate the Recall, Precision \& F1 as follows:\\
\begin{jumpin}
$Recall = \dfrac{True Positive}{True Positive + False Negative}$\vspace{0.1in}\\
$Precision = \dfrac{True Positive}{True Positive + False Positive}$\vspace{0.1in}\\
$F1 = \dfrac{2 * Precision * Recall}{Precision + Recall}$
\end{jumpin}
\subsubheader{Median \& average rank}
We use the median rank to indicate at which rank the 'standard' article would be ranked, based on its title or abstract. We do this by taking the median of the respective rank from each article. This gives us an indication of the behaviour of the articles in our validation set. This median rank (mostly) ignores the outliers, we therefore also use the average rank, which gives a more global indication, although this rank may be over-influenced by some outliers.
\subsubheader{Rank distribution}
To further analyse the ranking results, we plot the rank distribution to get an indication of the ranking-landscape.

\header{ Two-dimensional plot}
To create a plot, we transformed the 300-dimensional journal vectors into 2-dimensional vectors using TSNE based on pca\footnote{Principal component analysis}. These 2-dimensional vectors, representing the x \& y coordinate, can then be drawn in a plot. To visualize the preservation of journal-relatedness while converting the 300 dimensions to 2 dimensions, we create groupings using k-means. These groupings are created on the 300-dimensional vectors and are visualized in the plot using colors. We use the k-means groupings due to the lack of subject-based grouping values in our dataset.
\end{document}

>>>>

\documentclass[../../Thesis.tex]{subfiles}
\begin{document}
\header{Corpus}
The dataset we used for this research consists of articles published in 2017 which have been published in a journal that has, in 2017, atleast 150 publications. This results in a total dataset of $1.391.543$ articles from $3.759$ journals. Details about the corpus can be found in table~\ref{table:corpusSize}.
\begin{table}[hbt]
\begin{center}
\begin{tabular}{|l|c|c|c|}
\hline
 & Total count & Unique count & Average length \\
\hline\hline
Title words & $18.822.399$ & $939.665$ & 13,53  \\
\hline
Title tokens & $14.742.192$ & $230.805$ & 10,64 \\
\hline\hline
Abstract words & $264.653.020$ & $5.853.077$  & 190,19  \\
\hline
Abstract tokens & $171.474.473$ & $738.961$ & 123,71 \\
\hline\hline
Total words & $283.475.419$ & $6.209.769$  & 203,71 \\
\hline
Total tokens & $186.962.354$ & $763.475$ & 134,36 \\
\hline
\end{tabular}
\end{center}
\caption{Corpus size}\label{table:corpusSize}
\end{table}
\clearpage
\subheader{Text properties}
\citet{wiegand2018word} state that the Pareto distribution offers a good fit to the word occurrences in natural language. Their models show that, due to the additional parameters in the Pareto-III, the tail of the data fits better with the model than the Zipf model. This shows the relation between the word occurrences rank and the actual occurrences. This kind of word-occurrences distribution holds for many texts, including the writings of William Shakespeare and scientific texts and novels\cite{thurner2015understanding}.
The word occurrences in our corpus also follow the pattern of a Pareto distribution as described by~\citet{wiegand2018word}. Our distribution is visualized in figure~\ref{figure:wordTokenOccurrence}, which displays the occurrences of the first 500 tokens of the corpus.
\FloatBarrier
\begin{figure}[hbt]
\includegraphics[width=6.5in]{Plots/word_occurrences}
\caption{Word and token occurrences}\label{figure:wordTokenOccurrence}
\end{figure}
\FloatBarrier
\header{Datasets}
For this researched we used a (pre-made) tokenized dataset, created from the earlier described corpus, which reduces the total amount of words by 34\% (see table~\ref{table:corpusSize}). From this tokenized set, we created the embeddings and the TF-IDF feature vectors.
\subheader{Tokenization}
The following steps have been applied to the words to create a tokenized set:
\begin{enumerate}
\item{Removed punctuation}
\item{Removed all non-ASCII characters}
\item{Transformed all characters to lower-case}
\item{Removed stop-words, as provided by the NLTK\footnote{Natural Language ToolKit, \url{https://www.nltk.org/}} library}
\item{Removed numbers}
\item{Stemmed all words, using the stemmer provided by the NLTK library}
\end{enumerate}
These transformations reduced our dataset by 34\%, resulting in a tokenized set of $186.962.354$ tokens.

@@

\clearpage

@@

\header{Pipeline}
\subheader{Research environment}
The research will be done in a python-databricks environment, which uses spark, a library that offers tools to work with big-data. \citet{armbrust2015spark} state that Spark SQL lets programmers leverage the benefits of relational processing and lets SQL users call complex analytic libraries in Spark. This allows for much tighter integration between relational and procedural processing. The paper further states that Spark SQL makes it significantly simpler and more efficient to write data pipelines that mix relational and procedural processing while offering substantial speedups over previous SQL-on-Spark engines.\\
We procesessed the TF-IDF sets and the embedding sets via the same pipeline, using their common vector properties. This ensures comparable results, the pipeline is set-up as follows:
\begin{enumerate}
\item{Create training and validation set}
\item{Create journal embeddings}
\item{Categorize validation articles}
\item{Calculate performance metrics}
\end{enumerate}

@@

To calculate the distance between to vectors, cosine similarity is commonly used. We validated the quality of cosine similarity as a distance metrics by comparing it to other similarity matrices available in the SciPy\footnote{\url{https://www.scipy.org/}} library, which we used to calculate the distances. We calculate the similarities based on the normalized embeddings, and compared the distance metrics based on the default embedding set. Table~\ref{table:distanceMetrics} shows the results of this validation.\\

@@

\clearpage
\subheader{Performance measurement}
We use multiple metrics to validate the performance of the embedding sets and TF-IDF sets on the categorization task. These metrics are:
\begin{enumerate}
\item{F1-score}
\item{Median \& average rank}
\item{Rank distribution}
\end{enumerate}
\subsubheader{F1-score}
We define the positive \& negative metrics as follows:\\
\begin{jumpin}
$True Positive = $ Articles that are correctly matched to the current journal\\
$False Positive = $ Articles that are incorrectly matched to other journals\\
$False Negative = $ Articles that are incorrectly matched to the current journal\\
\end{jumpin}
We used these metrics to calculate the Recall, Precision \& F1 as follows:\\
\begin{jumpin}
$Recall = \dfrac{True Positive}{True Positive + False Negative}$\vspace{0.1in}\\
$Precision = \dfrac{True Positive}{True Positive + False Positive}$\vspace{0.1in}\\
$F1 = \dfrac{2 * Precision * Recall}{Precision + Recall}$
\end{jumpin}
\subsubheader{Median \& average rank}
We use the median rank to indicate at which rank the 'standard' article would be ranked, based on its title or abstract. We do this by taking the median of the respective rank from each article. This gives us an indication of the behaviour of the articles in our validation set. This median rank (mostly) ignores the outliers, we therefore also use the average rank, which gives a more global indication, although this rank may be over-influenced by some outliers.
\subsubheader{Rank distribution}
To further analyse the ranking results, we plot the rank distribution to get an indication of the ranking-landscape.
\subsubheader{Two-dimensional plot}
To create a plot, we transformed the 300-dimensional journal vectors into 2-dimensional vectors using TSNE based on pca\footnote{Principal component analysis}. These 2-dimensional vectors, representing the x \& y coordinate, can then be drawn in a plot. To visualize the preservation of journal-relatedness while converting the 300 dimensions to 2 dimensions, we create groupings using k-means. These groupings are created on the 300-dimensional vectors and are visualized in the plot using colors. We use the k-means groupings due to the lack of subject-based grouping values in our dataset.
\end{document}
